Resolvable Ambiguity

### 2019-07-10 paper/paper.tex
Updated the abstract and the intro.

\begin{proof}
  By defining two functions $\treetorun$ and $\runtotree$, the former from a tree to a run, the latter from run to tree, then showing these two functions to be inverses of each other.
\end{proof}

TODO: define a \emph{well-formed} linear encoding as a linear encoding without double parens. Ensure $\runtoword$ and $\lattice$ are defined.

\begin{lemma}
  All words $w \in L(A_{\posp{}})$ are well-formed linear encodings.
\end{lemma}

\begin{lemma}
  $\forall t \in L(T_D).\ \lattice(\runtoword(\treetorun(t))) = \lattice(t)$.
\end{lemma}

\begin{lemma}
  For all $t \in L(T_D)$ and every well-formed lattice encoding $w$ such that $\lattice(t) \subseteq \lattice(w)$, $w \in L(A'_{\posp{}})$.
\end{lemma}

\begin{lemma}
  We can define a function $\runprimetorun$ from runs in $A'_{\posp{}}$ to runs in $A_{\posp{}}$ such that $\forall R.\ \lattice(\runtoword(\runprimetorun(R))) \subseteq \lattice(\runtoword(R))$.
\end{lemma}

\begin{proof}
  By removing every transition that pushes or pops $\gamma$ and replacing every used transition $p \xrightarrow{'\pospl', +g} q$ that does not exist in $A_{\posp{}}$ with a transition $p \xrightarrow{'\reqpl', +g} q$.
\end{proof}

\begin{theorem}
  The existence of two distinct runs $R$ and $R'$, where $R$ is a successful run in $A_{\posp{}}$ and $R'$ in $A'_{\posp{}}$, and $\runtoword(R) = \runtoword(R')$, implies the existence of two distinct trees such that the lattice of one is entirely contained in the other.
\end{theorem}

\begin{proof}
  Assume $\runprimetorun(R') = R$. $\runprimetorun$ either leaves its argument unchanged, or changes the recognized word. If $R'$ and $R$ recognize the same word, and $\runprimetorun(R') = R$, then $R' = R$, but that is a contradiction.

  Since $\runprimetorun(R') \neq R$ we have two distinct trees $t_1 = \runtotree(\runprimetorun(R'))$ and $t_2 = \runtotree(R)$. Since $\lattice(\runtoword(\runprimetorun(R'))) \subseteq \lattice(\runtoword(R'))$ and $\lattice(\runtoword(R')) = \lattice(\runtoword(R))$ we have that $\lattice(t_1) \subseteq \lattice(t_2)$.
\end{proof}

@@

\begin{lemma}
  Given a $t \in L(T_D)$, we can construct a visibly pushdown automaton that accepts exactly $\words(t)$.
  \label{lemma:vpda-from-t}
\end{lemma}

@@

\begin{proof}
  Follows trivially from Lemma~\ref{lemma:vpda-from-t} and VPDAs being closed under union and difference.
\end{proof}

\begin{lemma}
\end{lemma}

>>>>

\begin{proof}
  (These are two attempts at showing the same thing, next attempt starts at ``Proof sketch''. I would strongly suggest reading ``Formal stuff'' first.)
  $A_{\posp{}}$ has a single initial state, and a single final state, call them $s$ and $f$ respectively. The two are distinct, and $f$ is reachable only through transitions that pop a stack symbol with $s$ and $f$ as the two first components. Transitions that push such symbols originate in $s$, and no other transitions originate in $s$. Thus all successful runs start in configuration $(s, \epsilon)$ and end in $(f, \epsilon)$, whereby every pushing transition in the run can be paired with a corresponding popping transition. We will now construct a tree $t \in L(T_D)$ in a bottom-up fashion.

  Select a subsequence of the run $R = (s, \epsilon) \; R_1 \; (p, s) \xrightarrow{a, +(p, q, w)} (p', (p, q, w) \cdot s) \; R_2 \; (q', (p, q, w) \cdot s \xrightarrow{b, -(p, q, w)} (q, s) \; R_2 \; (f, \epsilon)$ such that $R_2$ contains no pushing or popping operations. The only transitions that leave the stack unchanged in $A_{\posp{}}$ come from DFAs generated from productions in $D$. There is no way to transition between two such DFAs without using a pushing or popping transition, thus $R_2$ must be a successful run in the DFA for some labelled production $N -> l : r$ in $D$. $T_D$ must contain a corresponding production $N -> l(r)$. Now consider the sequence of input symbols $w'$ along the transitions in $R_2$. We have $w' \in L(r)$, since $R_2$ is a successful run in a DFA constructed from $r$. Thus $l(w') \in L_N(T_D)$. Now add parent nodes to this subtree for every label in $w$, e.g., if $w = l_1 \cdot l_2 \cdot l_3$ construct $l_1(l_2(l_3(l(w'))))$. We can then replace the subsequence of $R$ with this subtree, and then repeat the process until we have a tree with root $l'$, where $D$ has a labelled production $S -> l': r'$ for some $r'$ and the starting symbol $S$.

  To construct a successful run from a tree $t \in L(T_D)$ we can traverse the tree in a preorder, while at the same time constructing a run in $A_{\posp{}}$ starting from $(s, \epsilon)$. If the next terminal is in $\T$, i.e., a strictly zero-arity terminal, we take a transition originating in a DFA, which must be unique. If the next terminal is in $\Labels$, i.e., an unranked terminal, we take one of the added edges. However, this choice of edge might not be unique: there may be multiple edges from the current state to an appropriate DFA initial state. These stem from the original DFA having two transitions from the same state with the same label but different subscripts, e.g., $p \xrightarrow{l_\bot} q$ and $p \xrightarrow{l_\top} q'$. However, if both of these transitions were valid choices for the current tree that would imply that the DFA for that level can recognize two distinct words that differ only in subscripts. But distinctly subscripted terminals must originate from distinct occurances of marked non-terminals (since a $\top$ subscript corresponds to a mark, and a $\bot$ subscript corresponds to the absence of a mark), which would imply that the original regular expression is ambiguous if the marks are removed, which we explicitly ruled out in Section~\ref{sec:parse-time-disambiguation}.

% TODO: technically, we should perhaps mention that these two constructions are inverses of each other, but that's relatively easy to see once you understand them.
\end{proof}

\begin{lemma}
  The word recognized by a successful run in $A_{\posp{}}$ is the unique linear encoding of the lattice for the corresponding $t \in L(T_D)$.
\end{lemma}

\begin{lemma}
  For every successful run $R'$ in $A'_{\posp{}}$ there is one (or more) successful runs $R$ in $A_{\posp{}}$ such that the lattice encoded in the word recognized by $R'$ is larger than the lattice for the tree corresponding to $R$.
\end{lemma}

\begin{theorem}
  The existence of two distinct runs $R$ and $R'$, where $R$ is a successful run in $A_{\posp{}}$ and $R'$ in $A'_{\posp{}}$, implies the existence of two distinct trees such that the lattice of one is entirely contained in the other.
\end{theorem}

@@

\begin{lemma}
  Given a $t \in L(T_D)$, we can construct a visibly pushdown automaton that accepts exactly $\words(t)$.
\end{lemma}

@@
